\documentclass[a4paper,12pt]{article}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{url}
\usepackage{booktabs}

\newenvironment{cmds}
  {\begin{description}\renewcommand{\makelabel}[1]{\texttt{\textbackslash##1}\hfill}}
  {\end{description}}

\newenvironment{opts}
  {\begin{description}\renewcommand{\makelabel}[1]{\texttt{##1}\hfill}}
  {\end{description}}

\newenvironment{envs}
  {\begin{description}\renewcommand{\makelabel}[1]{\texttt{##1}\hfill}}
  {\end{description}}

\newcommand{\cmd}[1]{\texttt{\textbackslash#1}}
\newcommand{\opt}[1]{\texttt{#1}}
\newcommand{\env}[1]{\texttt{#1}}
\newcommand{\pack}[1]{\textit{#1}}
\newcommand{\argu}[1]{\textit{#1}}
\newcommand{\cls}[1]{\textsc{#1}}
\newcommand{\prog}[1]{\textbf{\textsf{#1}}}
\newcommand{\file}[1]{\texttt{#1}}

\author{Matthieu Weber\\ \texttt{mweber@mit.jyu.fi} \and Antti-Juhani Kaijanaho\\ \texttt{antkaij@mit.jyu.fi}}
\title{\cls{Gradu3} --- A Class for Writing Master's Thesis at the Faculty of
Information Technology of the University of Jyväskylä}

\begin{document}
\maketitle

\begin{abstract}
  This class is designed to facilitate the writing of a Master's
  Thesis (and other thesis) work at the Department of Information
  Technology of the University of Jyväskylä.  It may also be used
  elsewhere if it conforms to their requirements.
\end{abstract}

\section{Read this first}

The class is provided currently on a ''beta test'' basis.  We urge you
to consider it your honor-bound duty to \emph{let us know of any
deficiencies} that you may find when using this class.  First,
however, please verify that any strangeness you may witness is not
mandated by the university publishing unit specifications!

While this class has already been used to typeset several subsequently
approved theses successfully, \emph{it is always your responsibility,
as the author,} to ensure that the document meets the requirements
set by the university and the faculty in question. This class aids in
that, but it does not, and cannot, implement the all of the
typesetting requirements.

Please, subscribe to the mailing list Tutkielma-TeX
(\url{http://lists.jyu.fi/mailman/listinfo/tutkielma-tex}), and ask
there any questions you might have.  The subscribers there may be able
to help you faster than we can.

We may make changes to the class from time to time.  New releases will
be announced on that mailing list.

\section{Invocation}

Invoke the class with the usual \cmd{documentclass\{gradu3\}}.  The
options discussed in the following subsections should be given to the
\cmd{documentclass} command as optional arguments
(e.g. \cmd{documentclass[utf8,english]\{gradu3\}}).

\subsection{Language Options}

Languages options set hyphenation, titles translations and date formats to the
selected language. The available languages are:

\begin{opts}
\item[finnish] switches to Finnish language (default)
\item[english] switches to English language
\end{opts}

You can switch at any time from one of those languages to the other one with
\cmd{selectlanguage\{}\textit{language}\texttt{\}}.

\subsection{Input Character Set}

Input character set options allow to specify with what character set the source file
is written. Available character sets are:

\begin{opts}
\item[ascii] ASCII encoding (ISO 646)
\item[utf8] UTF-8 Unicode encoding (ISO 10646)
\item[latin1] ISO-8859-1 encoding (Western Europe languages) (default)
\item[latin2] ISO-8859-2 encoding (Central Europe languages)
\item[latin3] ISO-8859-3 encoding (Esperanto, Maltese)
\item[latin5] ISO-8859-5 encoding (Cyrillic)
\item[latin9] ISO-8859-15 encoding (Western European languages with the Euro symbol)
\item[applemac] Macintosh encoding
\item[ansinew] Windows 3.11 ANSI (extended ISO-8859-1) encoding
\item[cp1252] synonym for \opt{ansinew}
\item[cp1250] MS Windows 1250 (central and eastern Europe languages) code page
\item[decmulti] DEC Multinational Character Set encoding
\item[next] Next encoding
\item[cp437] IBM 437 code page
\item[cp437de] IBM 437 code page (German version)
\item[cp850] IBM 850 code page
\item[cp852] IBM 852 code page
\item[cp865] IBM 865 code page
\end{opts}

Nowadays, \opt{utf8} is almost always a safe choice, and it is the
default in \cls{gradu3}.

Note that these options are given to the \cmd{documentclass} command.
You should not invoke \cmd{usepackage\{inputenc\}} yourself!

\section{Other options}

\begin{opts}
\item[bachelor] This is a Bachelor's thesis.
\end{opts}

\section{Title and Metadata Commands}

Use the \cmd{maketitle} command after \cmd{begin\{document\}} but
before any of the actual text of the thesis.  The following commands
can (and in the case of mandatory commands, must) be used before the
\cmd{maketitle} command to specify the content of the title and
metadata pages.

\begin{cmds}
\item[title] document's title (mandatory)
\item[author] documents's author (mandatory).  This command can be
  repeated to specify more than one author.
\item[supervisor] thesis supervisor (mandatory).  This command can be
  repeated to specify more than one supervisor.
\item[date] document's publishing date (defaults to the current date)
  \textbf{deprecated, use \textbackslash setdate instead}
\item[setdate] takes three arguments: day, month, year (all numeric);
  sets document's publishing date (defaults to the current date)
\item[university] the university to which this thesis will be
  submitted (defaults to ``Jyväskylän yliopisto'' in Finnish and
  ``University of Jyväskylä'')
\item[department] the department  to which this thesis will be submitted (defaults to ``Tietotekniikan laitos'' in Finnish and ``Department of Mathematical Information Technology'' in English)
\item[subject] the formal subject (major or minor) whose thesis this is
  (defaults to ``Tietotekniikan'' in Finnish and ``in Information
  Technology'' in English
\item[studyline] study line (mandatory)
\item[tyyppi] type of work in Finnish (defaults to ``pro gradu -tutkielma'')
\item[type] type of work (defaults to ``pro gradu -tutkielma'' in
  Finnish and ``A Master's Thesis'' in English, except when the
  \opt{bachelor} option has been used)
\item[keywords] document's keywords in English (mandatory)
\item[avainsanat] document's keywords in Finnish (mandatory)
\item[contactinformation] author's contact information (mandatory). Can be
specified multiple times for multiple authors.
\item[tiivistelma] abstract in Finnish (mandatory)
\item[abstract] abstract in English (mandatory)
\item[translatedtitle] title in the other language (mandatory)
\end{cmds}

\section{Sectioning Commands}

The available sectioning commands are:
\begin{cmds}
\item[preface] Preface
\item[mainmatter] Marks the beginning of the main part of the document.
Should appear before the first \cmd{chapter}.
Inserts the table of contents, the list of tables and the list of figures.
Switches on chapter numbering; switches the page numbering from
lowercase roman to arabic, starting from 1 again.
\item[chapter] Beginning of a chapter
\item[section] Beginning of a section
\item[subsection] Beginning of a subsection
\item[appendix] starts the appendices.  Use \cmd{section} to give a
  title to each appendix.
\end{cmds}

Note that \cmd{subsubsection}, \cmd{paragraph} and \cmd{subparagraph} are
forbidden.

In addition to the previous commands, there are several useful
environments:

\begin{envs}
\item[thetermlist] is a list environment used to generate a term list;
  it generates a chapter heading before the list.
\item[chapterquote] allows you to add a quote to the beginning of a
  chapter.  It takes one mandatory argument: the attribution line for
  the quote.  You should start the text of the chapter immediately
  after this environment; do not leave empty lines there.
\end{envs}

\section{Citing sources}\label{sec:bibliographies}

The \cls{gradu3} class uses \cls{biblatex-chicago} for citations and
for the bibliography.  The following notes give the basics; see the
\textsc{Bib\LaTeX} and \cls{biblatex-chicago} manuals for more
advanced information.

You should list all your sources in one or more bibliographical
databases in the \textsc{Bib\TeX} format.  There are tools to help you
construct such databases, or you can do it by hand, if you want.

For each such database, put a command
\cmd{addbibresource\{\textit{file.bib}\}} in the document preamble
(that is, between the \cmd{documentclass} and \cmd{begin\{document\}}
commands), where \texttt{\textit{file.bib}} is the file name of the
particular database.

Add the command \cmd{printbibliography} just before the \cmd{appendix}
or \cmd{end\{document\}} command, whichever comes first.

To cite a source, there are several commands available.  Assuming that
there is a source labeled as \texttt{pierce02}, authored by Pierce and
published in 2002, then the command \cmd{cite\{pierce02\}} will
produce ``Pierce (2002)'' (without the quotes), and the command
\cmd{parencite\{pierce02\}} will produce ``(Pierce 2002)'' (without
the quotes).  To include a page reference, add it in square brackets
after the command name; for example,
\cmd{cite[p.\textasciitilde{}32]\{pierce02\}} will produce ``Pierce
(2002, p.~32)'' (without the quotes).

To countermand excessive cleverness on the part of \textsc{Bib\LaTeX},
if a citation command produces nothing or just an empty pair of
parentheses, put a \cmd{citereset\{\}} command somewhere between the
offending citation and the previous citation.

You will need to run the \texttt{biber} program in between \LaTeX{}
runs to complete the generation of the bibliograpgy.

\appendix
\section{Usage changes from \cls{gradu2} to \cls{gradu3}}

\begin{table}\centering
  \begin{tabular}{ll}
    \toprule
    \cls{gradu2}  & \cls{gradu3}  \\
    \midrule
    \opt{altlinespread} & behaviour not supported \\
    \opt{altsubsec}     & behaviour not supported \\
    \opt{copyright}     & behaviour not supported \\
    \opt{kandi}         & \opt{bachelor} \\
    \opt{logo}          & behaviour not supported \\
    \opt{lof}           & now default behaviour \\
    \opt{lot}           & now default behaviour \\
    \opt{newtitle}      & behaviour not supported  \\
    \opt{nocopyright}   & now default behaviour \\
    \opt{nonumbib}      & now default behaviour \\
    \opt{nottoc}        & now default behaviour \\
    \opt{numbib}        & behaviour not supported \\
    \opt{oldtitle}      & now default behaviour \\
    \opt{palatino}      & behaviour not supported \\
    \opt{shortthesis}   & \opt{bachelor} (for the typical use) \\
    \opt{surnamefirst}  & behaviour now unnecessary \\
    \opt{toc}           & behaviour not supported \\
    \bottomrule
  \end{tabular}
  \caption{Changes in option names from \cls{gradu2} to \cls{gradu3}}
  \label{tbl:optchange}
\end{table}

\begin{table}\centering
  \begin{tabular}{ll}
    \toprule
    \cls{gradu2}         & \cls{gradu3}  \\
    \midrule
    ---                  & \cmd{maketitle} \\
    ---                  & \cmd{supervisor} \\
    \cmd{acmccs}         & behaviour not supported \\
    \cmd{aine}           & \cmd{subject}\\
    \cmd{copyrightowner} & behaviour not supported \\
    \cmd{fulltitle}      & behaviour not supported \\
    \cmd{laitos}         & \cmd{department}\\
    \cmd{license}        & behaviour not supported \\
    \cmd{linja}          & \cmd{studyline}\\
    \cmd{paikka}         & behaviour not supported \\
    \cmd{setauthor}      & \cmd{author} (different syntax)\\
    \cmd{termlist}       & the \env{thetermlist} environment\\
    \cmd{tyyppi}         & \cmd{type}\\
    \cmd{yhteystiedot}   & \cmd{contactinformation}\\
    \cmd{yliopisto}      & \cmd{university}\\
    \cmd{ysa}            & behaviour not supported \\
    \bottomrule
  \end{tabular}
  \caption{Changes in command names from \cls{gradu2} to \cls{gradu3}}
  \label{tbl:cmdchange}
\end{table}

The old \cls{gradu2} class had a lot of options to cater for differing
tastes in the face of very underspecified departmental requirements.
Since \cls{gradu3} is based on new stricter requirements, there is no
longer need for so many options;\footnote{Remember, the more options a
  class has, the easier it is to break.} see
Table~\ref{tbl:optchange}.  Since the interface is being broken in
any case, changes have been made to commands
(Table~\ref{tbl:cmdchange}).

The \cls{gradu3} class uses the \pack{biblatex-chicago} package for
bibliography formatting.  This causes major changes to most documents.
Most notably, \env{thebibliography} and \cmd{bibliographystyle} cannot
be used, and a \cmd{bibliography} command should not be used.  See
Section~\ref{sec:bibliographies} for more information about how to use
the new bibliography tools.

Appendices cannot be \cmd{chapter}s in \cls{gradu3}; use \cmd{section}
to introduce each appendix after the \cmd{appendix} command.

\end{document}
